### 2015-09-09 vmcai-2016-tex/summary-heap.tex
Re-did the introduction, and tried to harmonize the rest of the paper somewhat.

%Our solution treats the heap as a hidden input parameter to every program.
%expressed completely in terms of a black-box symbolic input heap
%and the input reference parameters.
The collective state of all symbolic program reference values
is known as the \emph{symbolic heap}. Our symbolic heap 
is a field sensitive points-to graph with edges
annotated with disjunctive formulas describing aliasing
relationships. We represent the symbolic heap as a bipartite graph on
references, $\cfgnt{r}$, and constraint location pairs,
$\lp\phi\ \cfgnt{l}\rp$. The graph is expressed in the location map,
$\cfgnt{L}$, and the reference map, $\cfgnt{R}$. As done with the
environment, $\cfgnt{L}$ and $\cfgnt{R}$ are treated as partial
functions. $\cfgnt{L}(r) = \{(\phi\ \cfgnt{l})\ ...\}$ is the set
of constraint-location pairs (value set) in the heap associated with the given
reference. The constraint associated with each location is a guard
representing the conditions under which the associated reference
maps to the given location. By definition, any location appears 
at most once in any given value set. Unions of value sets containing
the same location are resolved by forming a disjunction of the constraints
from each set. $\cfgnt{R}(\cfgnt{l}, \cfgnt{f}) = \cfgnt{r}$ is the
reference associated with the given location-field pair in the
heap. References in the graph are immutable. Locations, however, are
mutated on updates.

\figref{fig:exampleHeap} shows an example symbolic heap. References are 
represented by circles and locations are represented by squares. Arrows leaving
references correspond to the value sets returned by the $\cfgnt{L}$ function, and 
arrows leaving the squares correspond to the $\cfgnt{R}$ function.

The soundness and completeness proofs in Section~\ref{sec:bisim} rely
on certain characteristics of the symbolic heap that are preserved by
the small-step semantics during symbolic execution:

\subsection{Symbolic Execution Semantics}

@@

A program, $\cfgnt{P}$, is a registry of classes, $\mu$, with
a tuple indicating a class, $\cfgnt{C}$, and a method, $\cfgnt{m}$,
where execution starts. For simplicity in presentation, booleans are the only primitive type, 
classes have only non-primitive fields, and methods have a single
parameter. Expressions, $\cfgnt{e}$, include statements, and they use
'$\cfgt{:=}$' to indicate assignment and '$\cfgt{=}$' to indicate
comparison.  The dot-operator for field access is replaced by
'$\cfgt{\$}$', and the dot-operator for method invocation is replaced
by '$\cfgt{@}$' since '.' is reserved in PLT Redex. There is no
explicit return statement; rather, the value of the last expression is
used as the return value. A variable is always indicated by \cfgnt{x}
and a value by \cfgnt{v}. A value can be a reference in the heap,
$\cfgnt{r}$, or any of the special values shown
in~\figref{fig:syntax}(a).  Only type correct programs are given as
input to the machine.

The state of the machine, $\cfgnt{s}$, includes the program
registry, $\mu$, a pair of functions defining the heap
$\lp\cfgnt{L}\ \cfgnt{R}\rp$, 
the path condition, $\phi$, a constraint defined over
references in the heap, the environment, $\eta$, for local variables,
and the continuation $\cfgnt{k}$. The registry never changes so it is
omitted from the state tuple in the rest of the presentation.  The
more complex structures, such as the environment and heap, are defined
as singly linked lists of mappings that start with empty, \cfgnt{mt}. These lists are treated as
partial functions so $\eta(\cfgnt{x}) = \cfgnt{r}$ is the reference
$\cfgnt{r}$ associated with variable $\cfgnt{x}$. The notation
$\eta^\prime = \eta[\cfgnt{x} \mapsto \cfgnt{v}]$ defines a new
partial function $\eta^\prime$ that is the same as $\eta$ except that
the variable $\cfgnt{x}$ now maps to $\cfgnt{v}$. The continuation
$\cfgnt{k}$ indicates with $\cfgnt{*}$ where the expression $\cfgnt{e}$ came from, stores
temporary computation, and keeps track of the next continuation. For
example, the continuation $\lp\cfgt{*}\ \cfgt{\$}\ \cfgnt{f}
\rightarrow \cfgnt{k}\rp$ indicates that the machine is evaluating the
expression for the object reference on which the field $\cfgnt{f}$ is
going to be accessed. Once the field access is complete, the machine
continues with $\cfgnt{k}$.

@@

The semantics are expressed as
rewrites on strings using pattern matching. Consider the rewrite rule
for the beginning of a field access instruction:
$$
\mprset{flushleft}
	\inferrule[Field Access(eval)]{}{
      \lp \cfgnt{L}\ \cfgnt{R}\ \phi\ \eta\ \lp \cfgnt{e}\ \cfgt{\$}\ \cfgnt{f}\rp \ \cfgnt{k}\rp  \rcom 
      \lp \cfgnt{L}\ \cfgnt{R}\ \phi\ \eta\ \cfgnt{e}\ \lp \cfgt{*}\ \cfgt{\$}\ \cfgnt{f} \rightarrow \cfgnt{k}\rp \rp 
	}
$$
If the string representing the current state matches the left side, then it
creates the new string on the right. In this example, the new string
on the right is now evaluating the expression $\cfgnt{e}$ in the field
access, and it includes the continuation indicating that it still
needs to complete the actual field access once the expression is
evaluated.

Other more complex rules, such as one to create a new instance of a
class, define constraints on the rewrites and more complex
transformations on the heap.
$$
\mprset{flushleft}
	\inferrule[New]{
      \cfgnt{r} = \mathrm{stack}_r\lp \rp\\
      l = \mathrm{fresh}_l\lp \cfgnt{C}\rp\\\\
      \cfgnt{R}^\prime = \cfgnt{R}[\forall \cfgnt{f} \in \mathit{fields}\lp \mathrm{C}\rp \ \lp \lp l\ \cfgnt{f}\rp  \mapsto \cfgnt{r}_\mathit{null} \rp ] \\\\
      \cfgnt{L}^\prime = \cfgnt{L}[\cfgnt{r} \mapsto \{\lp \cfgt{true}\ l\rp \}]
    }{
      \lp \cfgnt{L}\ \cfgnt{R}\ \phi\ \eta\ \lp \cfgt{new}\ \cfgnt{C}\rp \ \cfgnt{k}\rp  \rcom
      \lp \cfgnt{L}^\prime\ \cfgnt{R}^\prime\ \phi\ \eta\ \cfgnt{r}\ \cfgnt{k}\rp 
	}
$$ In this rule, when the string matches the new-expression, it is
        rewritten to use a new heap location where all of the fields
        for the new object point to $\cfgnt{r}_\mathit{null}$ and the
        location map points a new stack reference to that new object.
        New heap locations returned from $\mathrm{fresh}_l$ are monotonically
        increasing.
        Also note that references for objects from the new-expression are
        auxiliary literals, so they do not appear in any constraint in
        the symbolic heap as an aliasing option.

% \input{javalite-common}
%
The rewrite relations for the portion of the language
semantics that are common to both generalized symbolic execution and
the symbolic heap algorithm may be found in \cite{Hillery:2015}. The relation $\rcom$ includes all of these
rules. Excepting \textrm{N{\footnotesize EW}}, the rules do not update
the heap, and are largely concerned with argument evaluation in an
expected way. It is assumed that only type safe programs are input to
the machine so there is no type checking or error conditions. The
machine simply halts if no rewrite is enabled.

@@

>>>>

\subsection{Symbolic Execution}

@@

A program, $\cfgnt{P}$, is a registry of classes, $\mu$, with
a tuple indicating a class, $\cfgnt{C}$, and a method, $\cfgnt{m}$,
where execution starts. For simplicity in presentation, booleans are the only primitive type, 
classes have only non-primitive fields, and methods have a single
parameter. Expressions, $\cfgnt{e}$, include statements, and they use
'$\cfgt{:=}$' to indicate assignment and '$\cfgt{=}$' to indicate
comparison.  The dot-operator for field access is replaced by
'$\cfgt{\$}$', and the dot-operator for method invocation is replaced
by '$\cfgt{@}$'. There is no
explicit return statement; rather, the value of the last expression is
used as the return value. A variable is always indicated by \cfgnt{x}
and a value by \cfgnt{v}. A value can be a reference in the heap,
$\cfgnt{r}$, or any of the special values shown
in~\figref{fig:syntax}(a).  Only type correct programs are given as
input to the machine.

A \emph{symbolic heap} is a tuple $(\cfgnt{L}\ \cfgnt{R}\ \phi\ \eta)$, where the~\emph{location map} $\cfgnt{L}$,~\emph{reference map} $\cfgnt{R}$, and~\emph{environment} $\eta$ are partial functions, and the~\emph{path condition} $\phi$ is a predicate over the program variables. Given reference $r$, $\cfgnt{L}(r) = \{(\phi\ \cfgnt{l})\ ...\}$ is the set of constraint-location pairs (guarded value set) in the heap associated with $r$. The constraint  associated with each location is a guard
representing the conditions under which the associated reference
maps to the given location. By definition, any location appears 
at most once in any given value set. Unions of value sets containing
the same location are resolved by forming a disjunction of the constraints
from each set. $\cfgnt{R}(\cfgnt{l}, \cfgnt{f}) = \cfgnt{r}$ is the
reference associated with the given location-field pair in the
heap. The environment  $\eta$ maps local variables to references.

Conceptually, the symbolic heap may be thought of as a bipartite graph. \figref{fig:exampleHeap} shows an example symbolic heap graph. References are 
represented by circles and locations are represented by squares. Arrows leaving
references correspond to the value sets returned by the $\cfgnt{L}$ function, and 
arrows leaving the squares correspond to the $\cfgnt{R}$ function.

The machine state $\cfgnt{s}$ includes the program
registry $\mu$, the symbolic heap,
and the continuation $\cfgnt{k}$. The registry never changes so it is
omitted from the state tuple in the rest of the presentation.  The
more complex structures, such as the environment and heap, are defined
as singly linked lists of mappings that start with empty, \cfgnt{mt}. These lists are treated as
partial functions so $\eta(\cfgnt{x}) = \cfgnt{r}$ is the reference
$\cfgnt{r}$ associated with variable $\cfgnt{x}$. The notation
$\eta^\prime = \eta[\cfgnt{x} \mapsto \cfgnt{v}]$ defines a new
partial function $\eta^\prime$ that is the same as $\eta$ except that
the variable $\cfgnt{x}$ now maps to $\cfgnt{v}$. The continuation
$\cfgnt{k}$ indicates with $\cfgnt{*}$ where the expression $\cfgnt{e}$ came from, stores
temporary computation, and keeps track of the next continuation. For
example, the continuation $\lp\cfgt{*}\ \cfgt{\$}\ \cfgnt{f}
\rightarrow \cfgnt{k}\rp$ indicates that the machine is evaluating the
expression for the object reference on which the field $\cfgnt{f}$ is
going to be accessed. Once the field access is complete, the machine
continues with $\cfgnt{k}$.

@@

The semantics are expressed as
rewrites on strings using pattern matching. 
%Consider the rewrite rule
%for the beginning of a field access instruction:
%$$
%\mprset{flushleft}
%	\inferrule[Field Access(eval)]{}{
%      \lp \cfgnt{L}\ \cfgnt{R}\ \phi\ \eta\ \lp \cfgnt{e}\ \cfgt{\$}\ \cfgnt{f}\rp \ \cfgnt{k}\rp  \rcom 
%      \lp \cfgnt{L}\ \cfgnt{R}\ \phi\ \eta\ \cfgnt{e}\ \lp \cfgt{*}\ \cfgt{\$}\ \cfgnt{f} \rightarrow \cfgnt{k}\rp \rp 
%	}
%$$
%If the string representing the current state matches the left side, then it
%creates the new string on the right. In this example, the new string
%on the right is now evaluating the expression $\cfgnt{e}$ in the field
%access, and it includes the continuation indicating that it still
%needs to complete the actual field access once the expression is
%evaluated.
%
%Other more complex rules, such as one to create a new instance of a
%class, define constraints on the rewrites and more complex
%transformations on the heap.
%$$
%\mprset{flushleft}
%	\inferrule[New]{
%      \cfgnt{r} = \mathrm{stack}_r\lp \rp\\
%      l = \mathrm{fresh}_l\lp \cfgnt{C}\rp\\\\
%      \cfgnt{R}^\prime = \cfgnt{R}[\forall \cfgnt{f} \in \mathit{fields}\lp \mathrm{C}\rp \ \lp \lp l\ \cfgnt{f}\rp  \mapsto \cfgnt{r}_\mathit{null} \rp ] \\\\
%      \cfgnt{L}^\prime = \cfgnt{L}[\cfgnt{r} \mapsto \{\lp \cfgt{true}\ l\rp \}]
%    }{
%      \lp \cfgnt{L}\ \cfgnt{R}\ \phi\ \eta\ \lp \cfgt{new}\ \cfgnt{C}\rp \ \cfgnt{k}\rp  \rcom
%      \lp \cfgnt{L}^\prime\ \cfgnt{R}^\prime\ \phi\ \eta\ \cfgnt{r}\ \cfgnt{k}\rp 
%	}
%$$ In this rule, when the string matches the new-expression, it is
%        rewritten to use a new heap location where all of the fields
%        for the new object point to $\cfgnt{r}_\mathit{null}$ and the
%        location map points a new stack reference to that new object.
%        New heap locations returned from $\mathrm{fresh}_l$ are monotonically
%        increasing.
%        Also note that references for objects from the new-expression are
%        auxiliary literals, so they do not appear in any constraint in
%        the symbolic heap as an aliasing option. 
%        
% \input{javalite-common}
%
The rewrite relations for the portion of the language
semantics that are common to both generalized symbolic execution and
the symbolic heap algorithm may be found in \cite{Hillery:2015}. The relation $\rcom$ includes all of these
rules. Excepting \textrm{N{\footnotesize EW}}, the rules do not update
the heap, and are largely concerned with argument evaluation in an
expected way. It is assumed that only type safe programs are input to
the machine so there is no type checking or error conditions. The
machine simply halts if no rewrite is enabled.

@@

%Our solution treats the heap as a hidden input parameter to every program.
%expressed completely in terms of a black-box symbolic input heap
%and the input reference parameters.

Certain characteristics of the symbolic heap are preserved by
the small-step semantics during symbolic execution:
